Oregon 2013 Medication for Asthma Inhalers

### 2017-07-10 r_scripts/analysis/oregon_medication.Rmd
health analysis overall of lags

### Standard Regression Model

Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.

1. Quasipoisson

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

@@

@@

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

@@

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

@@

ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

2. Poisson

```{r regression model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

### Mixed models

1. Poisson

```{r mixed model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# populate outcome column
rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_mix_df[1,2:4] <- rr

rr_mix_df

plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
       width = 12, height = 8, units = "in")

@@

Remove the missing values for GWR smoke. 
But I use GWR smoke of zip for the case crossover study and add the county variable.

@@

## Distributed Lag Casecrossover Analysis

```{r overall, warning =F, echo = F, results='asis'} 
# dataframe list
method_list <- c('Geo-Weighted Smoke')

# dataframe to loop through
inhaler_ndc_lag1_df <- data.frame(inhaler_ndc) %>%
  filter(!is.na(geo_smk_pm_lag1_zip))

@@

# extract covariates from dataframe
covariates_df <- inhaler_ndc_lag1_df[, c(1:26, 74:79, 94:156)]

# extract pm values and divide by 10 and ordered
which(colnames(inhaler_ndc_lag1_df)=="geo_smk_pm_lag1_zip") # 136
pm_estimates_df <- inhaler_ndc_lag1_df[, 136]/10  # create 10 unit increases

@@

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  # filter(!is.na(wrf_smk_pm_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

@@

  # variable to model 
  var_name <- colnames(df_analysis[75])

  # set row number to fill
  row_n <- 1

  # only run the model if the dataframe has observations
  if(nrow(df_analysis) != 0){
  # conditional logistic regression model
  mod <- clogit(outcome ~ df_analysis[[75]] + wrf_temp_lag1_zip + strata(personkey), df_analysis)

  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list[row_n]
          point_estimates[row_n, 3] <- 0
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement

@@

@@

>>>>

### Standard Regression Model (Relative Risk)

1. Fixed-Quasipoisson
2. Fixed-Poisson
3. Mixed-Poisson

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")

@@

# methods
rr_df[,1] <- c("Fixed-Quasipoisson", "Fixed-Poisson", "Mixed-Poisson")

@@

# populate outcome column
rr_df[,2] <- factor(outcome_names, levels=unique(outcome_names))

## quasipoisson
smk_est_fq <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr_fq <- c(round(exp(smk_est_fq[1,2]),5),
         round(exp(smk_est_fq[1,2]-1.96*smk_est_fq[1,3]),5),
         round(exp(smk_est_fq[1,2]+1.96*smk_est_fq[1,3]),5))

rr_df[1,3:5] <- rr_fq

## poisson
smk_est_fp <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr_fp <- c(round(exp(smk_est_fp[1,2]),5),
         round(exp(smk_est_fp[1,2]-1.96*smk_est_fp[1,3]),5),
         round(exp(smk_est_fp[1,2]+1.96*smk_est_fp[1,3]),5))

rr_df[2,3:5] <- rr_fp

## mixed
smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr_mp <- c(round(exp(smk_est_mp[1,2]),5),
         round(exp(smk_est_mp[1,2]-1.96*smk_est_mp[1,3]),5),
         round(exp(smk_est_mp[1,2]+1.96*smk_est_mp[1,3]),5))

rr_df[3,3:5] <- rr_mp

@@

plot_rr <- ggplot(rr_df, aes(x=method, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

@@

ggsave("../plot_new/ndc_rr_geo.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

@@

Remove the missing values for GWR smoke. 
But I use GWR smoke of zip for the case crossover study.

@@

## Distributed Lag -- Casecrossover Analysis

### Lag

### 1. For Overall Plot

```{r overall lags, warning =F, echo = F, results='asis'} 
# dataframe list
method_list <- c('Geo-Weighted Smoke')

lag_list <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")

# dataframe to loop through
inhaler_ndc <- data.frame(inhaler_ndc) 
#  filter(!is.na(geo_smk_pm_lag1_zip))

# outcome name
outcome <- "Asthma Inhalers"

@@

# extract covariates from dataframe
covariates_df <- inhaler_ndc[, c(1:26, 74:79)] # 94:156 is lag

# extract pm values and divide by 10 and ordered
which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
temp_estimates_df <- inhaler_ndc[, 150:156]  # temperature

@@

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 7, ncol = 10, byrow = T)

colnames(point_estimates) <- c('outcome', 'pm_method',  'lag', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

# fill in the outcome namedataframe before method loop
point_estimates[, 1] <- outcome

for (i in 1:7){
geo_lag <- i + 32 
temp_lag <- i + 39

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df, temp_estimates_df)

df_analysis <- df_analysis %>% 
  # remove missing pm values
  filter(!is.na(df_analysis[geo_lag])) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

@@

  # variable to model 
  var_name <- colnames(df_analysis[geo_lag])

  # set row number to fill
  row_n <- i

  # only run the model if the dataframe has observations
  if(nrow(df_analysis) != 0){
  # conditional logistic regression model
  mod <- clogit(outcome ~ df_analysis[[geo_lag]] + df_analysis[[temp_lag]] + strata(personkey), df_analysis)

  # populate matrix
  point_estimates[row_n, 2] <- method_list
  point_estimates[row_n, 3] <- lag_list[row_n]
  point_estimates[row_n, 4] <- mod$n
  point_estimates[row_n, 5] <- mod$nevent
  # odds ratio
  point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 3] <- method_list[row_n]
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, 5] <- 0
          point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

}
# save point estimates as a dataframe
combined_point_est_df <- as_data_frame(point_estimates)

wrf_geo_lag <- combined_point_est_df %>% 
  select(3, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
         # row group by outcome
         rgroup = "Asthma Inhalers",
         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Lag", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_lag_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                             99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                             99, upper95)) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 10, 5:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
                              levels = unique(wrf_geo_lag_plot$outcome))

wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
                                levels = unique(wrf_geo_lag_plot$lag))

wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)

## ggplot 
  print_plot <- ggplot(wrf_geo_lag_plot,
    aes(x = lag, y = odds_ratio, color = lag), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Geo-smk Lag", 
                       values = c("red", "orange", "yellow", "green", "blue", "#32115C", "purple"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 7, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_health_lag.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

### 2. For Age Category

### 3. For Sex Category
